---
title: "Blocking Example"
output: html_notebook
---

```{r, message = FALSE}
library(tidyverse)
library(tidybayes)
library(generative)
review_theme()
```

* First, generate a design, specifying which people are given which treatments
and which control
* Then, write a function that simulates data given the design 
* Then write a function to estimate unknown effects given a design
* Loop that function over a few designs and many simulated datasets, to come up
with notion of power

First, we create the three different designs. The basic design just samples 90
people and assigns each to one of the three treatments at random. The blocked
design gives all three treatments to 30 people. The compromise randomly selects
2 treatments for each of 45 people.

```{r}
N <- 60
designs <- list(
  "basic" = basic_design(N),
  "blocked" = blocked_design(N),
  "compromise" = compromise_design(N)
)

designs
```

Next, we provide a function to generate data given one of these designs.

```{r}
effects <- design_effects(N, 3, 25)
x_blocked <- simulate(designs$blocked, effects)
x_df <- bind_cols(designs$blocked, x_blocked) %>%
  pivot_longer(-c("id", "treatment", "subject"), names_to = "feature")
```

Here are some of the effects visible from the blocked design. The first 10
features have true effects, the rest don't.

```{r, fig.height = 5, fig.width = 6}
plot_treatments(x_df) +
  geom_line(aes(group = subject), alpha = 0.4, size = 0.2) 
```

This is the comparable figure if we didn't have any blocking. Note that the
feature-level treatment effects are exactly the same, but now we can't draw
lines linking the same individuals. The effects for the first 10 features are
visible, but in some cases less obvious than when blocking was used.

```{r, fig.height = 5, fig.width = 6}
x_basic <- simulate(designs$basic, effects)
x_df <- bind_cols(designs$basic, x_basic) %>%
  pivot_longer(-c("id", "treatment", "subject"), names_to = "feature")
plot_treatments(x_df)
```

Now we'll write a small program to estimate the posterior on betas.

```{r, messgae = FALSE}
fit <- multinomial_logistic(x_blocked, designs$blocked)
treatment_estimates <- list()
treatment_estimates[["blocked"]] <- process_estimate(fit, "treatment")
truth <- process_truth(effects$treatment)
```

```{r}
comparison <- treatment_estimates$blocked %>%
  group_by(feature, treatment) %>%
  summarise(posterior_mean = mean(value)) %>%
  left_join(truth)

ggplot(comparison) +
  geom_abline(slope = 1) +
  geom_point(aes(value, posterior_mean))
```

```{r, message = FALSE}
fit <- multinomial_logistic(x_basic, designs$basic)
treatment_estimates[["basic"]] <- process_estimate(fit, "treatment")
```
```{r}
bind_rows(treatment_estimates, .id = "design") %>%
  mutate(design = fct_recode(design, unblocked = "basic")) %>%
  ggplot(aes(feature, value, col = as.factor(treatment))) +
  stat_dots() +
  scale_color_brewer(palette = "Set2") +
  geom_point(data = truth, size = 1.6, col = "#0c0c0c") +
  geom_point(data = truth, size = 0.8) +
  labs(x = "Taxon", y = "Effect", col = "Treatment") +
  facet_grid(design ~ .)

ggsave("~/Downloads/blocking_counts.png", dpi = 500, height = 4, width = 7)
```

The relative efficiency of the blocked vs. unblocked estimator (for this setting
of $N$, $\sigma^2_{\text{treatment}}$, $\sigma^2_{\text{subject}}$) is about
1.35.

```{r}
bind_rows(treatment_estimates, .id = "design") %>%
  group_by(design, feature, treatment) %>%
  summarise(sd = sd(value)) %>%
  group_by(design) %>%
  summarise(mean(sd))
```
