---
title: "R Notebook"
output: html_notebook
---

```{r}
library(reticulate)
library(tidyverse)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 13),
    legend.position = "bottom"
  )
theme_set(th)
```

The block below is needed to install the simulator. In theory, it should be
possible to install using reticulate, but I find it easier to just use the
terminal.

```{sh, eval = FALSE}
conda create -n covid
conda activate covid
conda install numpy=1.21
pip install covasim
pip install sciris
pip install optuna
```

```{r}
use_condaenv("covid", required = TRUE)
```

```{python}
import covasim as cv

# a few kinds of interventions
testing = cv.test_prob(symp_prob=0.8, asymp_prob=0.001)
tracing = cv.contact_tracing(trace_probs=0.1, trace_time=5.0)
distancing = cv.change_beta(changes=1.0, days=10)
```

This is a calibration example.

```{python}
import sciris as sc
import covasim as cv

# Create default simulation
pars = sc.objdict(
  pop_size = 1e4,
  start_day = "2020-02-01",
  end_day = "2020-04-11",
  beta = 0.015,
  rel_death_prob = 1.0,
  interventions = cv.test_num(daily_tests="data")
)
sim = cv.Sim(pars=pars, datafile="calibration_data.csv")

# Parameters to calibrate -- format is best, low, high
calib_pars = {
  "beta": [pars.beta, 0.005, 0.20],
  "rel_death_prob": [pars.rel_death_prob, 0.5, 3.0]
}

# Run the calibration
n_trials = 25
n_workers = 4
calib = sim.calibrate(calib_pars=calib_pars, n_trials=n_trials, n_workers=n_workers)
```

Now we can use these calibration parameters.

```{python}
pars.beta = calib.best_pars["beta"]
pars.rel_death_prob = calib.best_pars["rel_death_prob"]
```

```{python}
```

Now that we have a reasonably calibrated model, let's try to learn an emulator
that allows us to see the effect of tracing and testing measures.
```{python}
import numpy as np
np.linspace(0.1, 1.0, 10)
```

```{python}
import numpy

results = []

for symp_prob in np.linspace(0, 1, 5):
  for trace_prob in np.linspace(0, 1, 5):
    for trace_time in np.linspace(0, 5, 4):
      testing = cv.test_prob(symp_prob=symp_prob)
      tracing = cv.contact_tracing(trace_probs=trace_prob, trace_time=trace_time)
      pars.interventions = [testing, tracing]
      
      for rep in range(2):
        cv.Sim(pars=pars)
        sim.run()
        
        results.append({
          "symp_prob": symp_prob,
          "trace_prob": trace_prob,
          "trace_time": trace_time,
          "results": sim.to_df()
        }
```

```{r}
py$results


#ggplot(py$sim_res) +
#  geom_point(aes(t, new_infections))
```

